NLIZE: a Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-24 paper/background.tex
update background

The inference task can be quite challenging, considering the ambiguousness of natural language (especially from learning algorithm's point of view). Take the following sentences as an example (here \textbf{P} refer to premise, \textbf{H} refer to hypothesis):  (\textbf{P}) Facebook's IPO electrified the general public; (\textbf{H1}) Facebook went public; (\textbf{H2}) General Electric went public; (\textbf{H3}) People ignored Facebook's IPO. The literal semantic similarity between ``Electric'' and ``electrified'' may trick the model to predict \textbf{H2} as \emph{entailment}. The model likely will also fail to understand the link between ``went public'' and ``IPO'', therefore, mistake \textbf{H1} as \emph{neutral}.

%\shusen{How does the NLI related to the other challenges in NLP}
The grand challenge in Natural Language Processing (NLP) is to have machine to acquire
deep comprehension of textual information. Major tasks include Natural Language Inference,
Machine Translation, Questions Answering, Summarization, and etc.
Recently, neural networks models have performed strongly on these tasks.
%
Recently ~\cite{BowmanAngeliPotts2015} proposed a large dataset for this task. We will examine
the decomposable attention model~\cite{parikh2016emnlp} on this dataset with perturbation.

@@

\subsection{Neural Network Model in NLP}
\shusen{neural linguistic models?}

\shusen{What is the basic understanding of neural network What is the end-to-end}

% \subsection{Interprebilty Challenges}
Compared with Conputer Vision, the discrete nature of words and sentences
presents additional challenge for
interpreting the model, since many visualization technique often employed
for images rely on continuous nature of the input space (e.g., one can interpolate
real values much easier than interplolate between words/sentences).
%
On the other hand, a majority of recent NLP neural networks share the nature of
end-to-end model, where the entire model operates as a black-box that takes
vectorized input and yields final prediction for a specific task.
Therefore, to study the inside of end-to-end neural models, it is important to
explore the interaction between intermediate representations and predictions.

\subsection{Attention Mechansim}
\label{sec:attention}
The introduction of attention mechanism~\cite{bahdanau2014neural} allows
pairwise interaction between hidden states. This interaction can be naturally explained
as a form of alignment which exposes an interpretable layer in end-to-end neural networks.
Recently attention has contributed to many strongly-performed NLP models
~\cite{parikh2016emnlp},~\cite{rush2015neural},~\cite{yang2016hierarchical},
~\cite{seo2016bidirectional},~\cite{schwartz2017high}.
We will focus on its application in Natural Language Inference task as proposed in~\cite{parikh2016emnlp}.
However notice that our visualization can be generalized to other attention models as well.

>>>>

Considering the ambiguousness  of natural language and polysemy words, the inference task can be quite challenging (especially from learning algorithm's point of view). Take the following sentences as an example (here \textbf{P} refer to premise, \textbf{H} refer to hypothesis):  (\textbf{P}) Facebook's IPO electrified the general public; (\textbf{H1}) Facebook went public; (\textbf{H2}) General Electric went public; (\textbf{H3}) People ignored Facebook's IPO. The literal semantic similarity between ``Electric'' and ``electrified'' may trick the model to predict \textbf{H2} as \emph{entailment}. The model likely will also fail to understand the link between ``went public'' and ``IPO'', therefore, mistake \textbf{H1} as \emph{neutral}.
%
At a first glance, the task of natural language inference may seem less practical compared to other common NLP challenges such as machine translation, however, the ability to distinguish entailment and contradiction relationship is fundamental to understanding natural language at large. 
%The grand challenge in Natural Language Processing (NLP) is to have machine to acquire
%deep comprehension of textual information. Major tasks include Natural Language Inference,
%Machine Translation, Questions Answering, Summarization, and etc.
%Recently, neural networks models have performed strongly on these tasks.
%
Recently, Bowman et al. introduce a large corpus~\cite{BowmanAngeliPotts2015} for the NLI task, which helps spawn a new wave of effective neural network based models. In this work, we focus on the analysis of the decomposable attention model~\cite{parikh2016emnlp} on this dataset.

@@

\subsection{Neural Network Model in NLP}
%\shusen{neural linguistic models?}
%\shusen{What is the basic understanding of neural network What is the end-to-end}
There are many connection and distinction between neural network model employed for vision tasks and NLP tasks.
On the one hand, as discussed in the introduction, the discrete nature of words and sentences presents additional challenge for
interpreting the NLP model.
%
On the other hand, a majority of recent NLP neural networks share the nature of
end-to-end model, where the entire model operates as a black-box that takes
vectorized input and yields final prediction for a specific task.
Therefore, to make sense of how predictions are made inside of end-to-end neural models, it is important to
explore the interaction between intermediate representations and predictions.

Many of the recent NLP end-to-end neural network models, despite having potentially drastic different network architecture, share a similar high-level design that consist of three distinct stages (encode, attent/align, classify, see Fig.~\ref{fig:modelPipeline}). 
These models usually take pre-trained word vectors (a numerical vector representation of individual words, where the semantic similarities are encode by distances in the high-dimensional space~\cite{MikolovSutskeverChen2013, PenningtonSocherManning2014}) as input, the

\subsection{Attention Mechanism}
\label{sec:attention}
The introduction of attention mechanism~\cite{bahdanau2014neural} allows
pairwise interaction between hidden states. This interaction can be naturally explained
as a form of alignment which exposes an interpretable layer in end-to-end neural networks.
Recently attention has contributed to many strongly-performed NLP models
~\cite{parikh2016emnlp},~\cite{rush2015neural},~\cite{yang2016hierarchical},
~\cite{seo2016bidirectional},~\cite{schwartz2017high}.
We will focus on its application in Natural Language Inference task as proposed in~\cite{parikh2016emnlp}.
However notice that our visualization can be generalized to other attention models as well.
